### 2022-12-28 appendix.Rmd
added appendix with figures

#

```{r Figure2, fig.width = 8, fig.height = 10, fig.cap = "Between-group discrepancies for all evaluation criteria split by *M*."}
### For the appendix -- other splits for the main results as plot (N is most important)
# Plot the results, by M
ldf %>% 
  group_by(method, Objective, M) %>% 
  summarise(Mean = mean(value)) %>% 
  filter(method != "random") %>% 
  mutate(
    Objective = ordered(
      Objective, 
      levels = c("means", "sd", "skew", "kur", "cor"),
      labels = c("M", "SD", "Skew", "Kurtosis", "Correlation"))
  ) %>% 
  ggplot(aes(x = M, y = Mean, colour = method)) + 
  geom_line(aes(linetype = method), size = .85) + 
  facet_grid(rows = vars(Objective), scales = "free") + 
  ylab("Mean discrepancy") +
  theme_bw(base_size = 16) + 
  scale_x_continuous(breaks = 2:5)
```

```{r Figure3, fig.width = 8, fig.height = 10, fig.cap = "Between-group discrepancies for all evaluation criteria split by *SD*."}
# Plot the results, by SD
ldf %>% 
  group_by(method, Objective, SD) %>% 
  summarise(Mean = mean(value)) %>% 
  filter(method != "random") %>% 
  mutate(
    Objective = ordered(
      Objective, 
      levels = c("means", "sd", "skew", "kur", "cor"),
      labels = c("M", "SD", "Skew", "Kurtosis", "Correlation"))
  ) %>% 
  ggplot(aes(x = SD, y = Mean, colour = method)) + 
  geom_line(aes(linetype = method), size = .85) + 
  facet_grid(rows = vars(Objective), scales = "free") + 
  ylab("Mean discrepancy") +
  theme_bw(base_size = 16) + 
  scale_x_continuous(breaks = 1:3)
```

```{r Figure4, fig.width = 8, fig.height = 10, fig.cap = "Between-group discrepancies for all evaluation criteria split by *K*."}
# Plot the results, by K
ldf %>% 
  group_by(method, Objective, K) %>% 
  summarise(Mean = mean(value)) %>% 
  filter(method != "random") %>% 
  mutate(
    Objective = ordered(
      Objective, 
      levels = c("means", "sd", "skew", "kur", "cor"),
      labels = c("M", "SD", "Skew", "Kurtosis", "Correlation"))
  ) %>% 
  ggplot(aes(x = K, y = Mean, colour = method)) + 
  geom_line(aes(linetype = method), size = .85) + 
  facet_grid(rows = vars(Objective), scales = "free") + 
  ylab("Mean discrepancy") +
  theme_bw(base_size = 16) + 
  scale_x_continuous(breaks = 2:4)
```

```{r Figure5, fig.width = 8, fig.height = 10, fig.cap = "Between-group discrepancies for all evaluation criteria split by *r*."}
# Plot the results, by r
ldf %>% 
  group_by(method, Objective, r) %>% 
  summarise(Mean = mean(value)) %>% 
  filter(method != "random") %>% 
  mutate(
    Objective = ordered(
      Objective, 
      levels = c("means", "sd", "skew", "kur", "cor"),
      labels = c("M", "SD", "Skew", "Kurtosis", "Correlation"))
  ) %>% 
  ggplot(aes(x = r, y = Mean, colour = method)) + 
  geom_line(aes(linetype = method), size = .85) + 
  facet_grid(rows = vars(Objective), scales = "free") + 
  ylab("Mean discrepancy") +
  theme_bw(base_size = 16) + 
  scale_x_continuous(breaks = 0:5 / 10)
```