### 2024-03-11 paper.Rmd
text

---
title             : "Extending the Bicriterion Approach for Anticlustering: Optimal and Heuristic Approaches"
shorttitle        : "Extending the Bicriterion Approach for Anticlustering"

author:
  - name          : "Martin Papenberg"
    affiliation   : "1"
    corresponding : yes
    address       : "Heinrich-Heine-Universität Düsseldorf, Institut für Experimentelle Psychologie, Universitätsstraße 1, 40225 Düsseldorf, Germany"
    email         : "martin.papenberg@hhu.de"
  - name          : "Martin Breuer"
    affiliation   : "1"
    corresponding : no
  - name          : "Max Diekhoff"
    affiliation   : "1"
    corresponding : no
  - name          : "Gunnar W. Klau"
    affiliation   : "1"
    corresponding : no

affiliation:
  - id            : "1"
    institution   : "Heinrich Heine University Düsseldorf"

authornote: |

  Martin Papenberg, Department of Experimental Psychology, Heinrich 
  Heine University Düsseldorf. Martin Breuer, Gunnar W. Klau, Max Diekhoff, 
  Department of Computer Science, Heinrich Heine University Düsseldorf.

abstract: |

    Numerous applications in psychological research require that a data set is partitioned via the maximization---instead of the more common minimization---of a clustering criterion. Such anticlustering seeks for high similarity between groups (i.e., maximum diversity) or high pairwise dissimilarity within groups (i.e., maximum dispersion). Brusco et al. (2020) proposed a powerful bicriterion algorithm (BILS) that simultaneously seeks for maximum diversity and dispersion, introducing the bicriterion approach for anticlustering. Here, we advance the bicriterion approach in several ways. As our main contribution, we combine a novel exact algorithm for maximum dispersion with the BILS heuristic to obtain high diversity on top of an optimal dispersion. Despite its theoretical computational hardness, our optimal method scales to rather large data sets. It generates multiple partitions that we use as initializations of the BILS's local search, thus guaranteeing that the BILS also returns a partition with optimal dispersion. In a simulation study, we compared several adaptations of the BILS. To obtain high diversity, initializing the BILS with multiple optimal partitions is preferable to using only one. Moreover, in data sets where maintaining optimal dispersion severely restricts the search for feasible partitions, the iterated local search phase of the BILS is crucial to obtain high diversity. While the original BILS oftentimes finds optimal dispersion values, our simulation illustrates conditions when it does not. The optimal dispersion algorithm and the BILS including our extensions are available via the free and open source R package `anticlust`. Practical examples are included to illustrate their application.

bibliography      : ["lit.bib"]

keywords          : "Anticlustering, maximum dispersion, optimal algorithm, bicriterion optimization"
wordcount         : "12345"

floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no

output            : 
  papaja::apa6_pdf

header-includes:
  - \usepackage{setspace}
  - \usepackage{float}
  - \usepackage{amsmath}
  - \usepackage{hyperref}
  - \floatstyle{plaintop}
  - \restylefloat{figure}
  - |
    \makeatletter
    \renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
      {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
      {-1em}%
      {\normalfont\normalsize\bfseries\typesectitle}}

    \renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
      {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
      {-\z@\relax}%
      {\normalfont\normalsize\bfseries\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
    \makeatother
  - |
    \makeatletter
    \renewcommand*\env@matrix[1][*\c@MaxMatrixCols c]{%
      \hskip -\arraycolsep
      \let\@ifnextchar\new@ifnextchar
      \array{#1}}
    \makeatother

csl               : "`r system.file('rmd', 'apa7.csl', package = 'papaja')`"
documentclass     : "apa7"

---

```{r setup, include = FALSE}
library(papaja) # use the development version: `remotes::install_github("crsh/papaja@devel")`; currently 0.1.1.9001
library(anticlust)
library(knitr)
# library(dplyr)
# library(ggplot2)
# library(tidyr)
# library(DescTools)
# for formatting numbers in Rmarkdown, use my package prmisc (https://github.com/m-Py/prmisc)
library(prmisc) # remotes::install_github("m-Py/prmisc") 
options(digits = 2)
knitr::opts_chunk$set(message=FALSE, warning = FALSE, echo = FALSE, dev = "cairo_pdf")

```

Partitioning data sets via anticlustering has uses in many research fields including psychology [@brusco2019; @schaper2023], test assessment [@gierl2017], education [@krauss2013; @baker2002], artificial intelligence [@steghofer2013], machine learning [@mauri2023], network systems [@mohebi2022], and operations research [e.g. @gallego2013; @gliesch2021]. Anticlustering seeks for similarity between groups and heterogeneity within groups by maximizing a clustering objective criterion. Thus, anticlustering is the logical and mathematical opposite of classical clustering, which partitions data sets via minimization of the same criteria [@steinley2006; @brusco2019; @spath1986].

Formally, anticlustering is used to partition $N$ objects into $K$ groups $C = (C_1, \ldots, C_K)$, where $C_k$ is the set of $n_k$ objects in the $k^{th}$ group, and $K$ and $n_k$ are predetermined by the application. In many anticlustering applications, equal-sized groups are required, i.e., $n_k = \frac{N}{K}$ [@schulz2022; @gallego2013]. The partitioning is characterized by assignment to exhaustive and mutually exclusive subgroups:

\begin{align}
\bigcup\limits_{k = 1}^{K} C_k = X \label{formalization:1} \\
C_j \cap C_k = \emptyset, \; \forall j, k \in \{1, \ldots, K\}, \; j \ne k\label{formalization:2} \\
\vert C_k \vert = n_k, \; \forall k \in \{1, \ldots, K\} \label{formalization:3}
\end{align}

We assume that a data matrix $\mathbf{Z}$ is used as input for computing the anticlustering criterion. $\mathbf{Z}$ is either a feature matrix $\mathbf{X}_{N \times M} = \{x_{ij}\}_{N \times M}$ where the $M$ columns contain numeric measurements on the $N$ objects, or a symmetric matrix $\mathbf{D}_{N \times N} = \{d_{ij}\}_{N \times N}$ of pairwise dissimilarity measurements. The partitioning is obtained by maximizing a clustering criterion $f_i(C, \mathbf{Z})$. For example, *k*-means anticlustering aims to maximize between-group similarity by maximizing the error sum of squares, i.e., the sum the squared distances between each object and its cluster center [@spath1986; @papenberg2024]:

$$
f_1(C, \mathbf{X}_{N \times M}) =
  \sum\limits_{j=1}^{M} 
  \sum\limits_{k=1}^{K} 
  \sum\limits_{i \in C_k}
  (x_{ij} - \overline{x}_j ^{(k)})^2
$$

The $k^{th}$ cluster center $\overline{\mathbf{x}}^{(k)} = (\overline{x}_1^{(k)}, \overline{x}_2^{(k)} \ldots, \overline{x}_P^{(k)})$ contains the mean values of the $M$ variables computed across all observations belonging to the  $k^{th}$ group:

$$
\overline{x}_j ^{(k)} = \frac{1}{n_k} \sum\limits_{i \in C_k} x_{ij}
$$

Maximizing $f_1$ ensures that the cluster centers, i.e., the mean values of the $M$ measurements, are as close to each other as possible [@spath1986]. The *k*-plus criterion is an extension of the $k$-means criterion for anticlustering that can be implemented by adding fictitious to the data matrix $\mathbf{X}_{N \times M}$ [@papenberg2024]. *K*-plus anticlustering is used to equate not only means between clusters, but higher order moments such as the variance or skewness as well. Because the approach can be reduced to an augmentation of the data set, *k*-plus anticlustering leaves the original *k*-means criterion $f_1$ unchanged. For example, to equate means and variances, we define a new data matrix $\mathbf{Y}_{N \times P} = \{y_{ij}\}_{N \times P}$ with $y_{ij} = (x_{ij} - \overline{x}_j)^2$ and augment the original matrix with it. Then, the *k*-plus criterion is computed as

$$
f_2(C, \mathbf{(X | Y)_{N \times 2M}}) =
  \sum\limits_{j=1}^{M} 
  \sum\limits_{k=1}^{K} 
  \sum\limits_{i \in C_k}
  (x_{ij} - \overline{x}_j ^{(k)})^2 + (y_{ij} - \overline{y}_j ^{(k)})^2
$$

Another important anticlustering criterion is the *diversity*, which uses a dissimilarity matrix $\mathbf{D}_{N \times N}$ as input. It is computed as the within-group sum of pairwise dissimilarities across all clusters [@brusco2019; @papenberg2020; @feo1990]:

$$
f_3(C, \mathbf{D}_{N \times N}) = \sum_{(i < j) \in C_k} d_{ij}
$$

Maximizing the diversity simultaneously minimizes the sum of distances between objects not in the same group. Therefore, the diversity represents within-group heterogeneity as well as between-group similarity [@feo1990]. Being a sum instead of an average, however, it no longer adequately captures between-group similarity if group sizes are not equal. The diversity is oftentimes obtained by first converting the data matrix $\mathbf{X}_{N \times M}$ into a matrix of pairwise Euclidean distances. However, most anticlustering algorithms that use a dissimilarity matrix as input are agnostic as to how it was was generated [@brusco2019; cf. @schulz2022]. If $\mathbf{D}_{N \times N}$ is computed via the squared Euclidean distances, we note an important equivalence between the k-means criterion and the diversity. The k-means criterion can also be expressed using the pairwise squared Euclidean distances between data points (Brusco, 2006; Späth, 1980). Based on $\mathbf{D}_{N \times N}$, the k-means criterion is then computed as

$$
f_1(C, \mathbf{X}_{N \times M}) =
f_4(C, \mathbf{D}_{N \times N}) = \sum\limits_{k = 1}^{K}\left( \frac{1}{n_k} \sum_{(i<j)\in C_k} d_{ij} \right)
$$

That is, "the sum of squared distances of a collection of points from the centroid associated with those points is equal to the sum of the pairwise squared distances between those points divided by the size (number of objects) of the collection" (Brusco, 2006, p. 350).

An important special case of anticlustering applications is that all groups are equal-sized, i.e., $n_k = \frac{N}{K}$ ($k = 1, \dots K$). Here, we obtain

$$
\mathit{SSE} = \frac{N}{K} \sum\limits_{k = 1}^{K} \sum_{(i<j)\in C_k} d_{ij}
$$

In this case, the normalizing factor $\frac{N}{K}$ can be ignored when maximizing $f_1$. Thus, the criterion reduces to the sum of the (squared) distances between elements in the same cluster:

$$
\mathit{SSE_{Diversity}} = \sum\limits_{k = 1}^{K}\sum_{(i<j)\in C_k} d_{ij}
$$

This term corresponds to the standard diversity $f_2$, with the distances $d_{ij}$ representing squared Euclidean distances. We denote $f_4$ as the *average diversity criterion*, which maintain unconditional equivalence between k-means and the diversity.

Unlike $f_1$, $f_2$, $f_3$, and $f_4$, the *dispersion* criterion does not represent between-group similarity, but is a pure measure of within-group heterogeneity. It is defined as the minimum dissimilarity across any two objects that are part of the same group [@brusco2019; @fernandez2013]:

$$
f_3(C) = \textrm{min}_{(i < j) \in C_k} d_{ij}
$$

The dispersion is a measure of the worst-case pairwise dissimilarity. Maximizing the dispersion ensures that any two objects in the same group are as dissimilar from each other as possible.

@brusco2019 outlined the importance of simultaneously considering multiple criteria when tackling anticlustering problems. In particular, they argued that anticlustering algorithms should incorporate an objective of between-group similarity---such as the diversity---and pairwise within-group dissimilarity---i.e., the dispersion---at the same time.

## Algorithms for anticlustering

Heuristics dominate (local maximum search @spath1986; @weitz1998, multiple repetitions robust for k-means @steinley2007, ILS @brusco2017, VARIABLE NEIGHBORHOOD SEARCH @urovsevic2014; tabu search @gallego2013; multiple improvement phases based on initial local maximum search @yang2022). Exact algorithms for diversity [@papenberg2020; @schulz2022] and for the dispersion [@fernandez2013; @gliesch2021; @brucker1978]. Exact algorithms for diversity perform much worse than for the reversed clustering problems, where many hundred elements can often be processed. Maximum diversity can be achieved with much better success. Our novel contribution is to combine optimal maximum dispersion with high diversity. We present a novel algorithm for maximum diversity that is similar to the approaches by @fernandez2013 and @gliesch2021 and is based on the logic that was already introduced by @brucker1978. It scales to hundreds or even thousands of data points and therefore can be used to solve most real-world applications we have in mind. We then proceed to discuss how on top of an optimal dispersion, we can implement high diversity. A logical next step is to obtain optimal maximum diversity as well, which is however only limited to rather small data sets, as we show. Due to its bicriterion nature, the BILS by @brusco2019 is naturally suited to maintain optimum dispersion while optimizing diversity. We discuss several adaptations of the BILS, which are compared in a simulation study.

In within-subjects experiments, participants are presented with several stimulus sets that are affiliated with the different experimental conditions. Due to carry-over effects, showing the same stimuli in different conditions is oftentimes prohibited. This is evident in studies on memory, where studying a list of words in one conditions necessarily makes it impossible to repeat their study without being influenced by the earlier study phase; memory cannot be erased after an experimental block has been completed [@lahl2006]. While the experimental logic dictates that different lists must be used, the lists should be as parallel as possible with regard to variables that affect the response variable. @lintz2021 gave two reasons as to why it is important that study participants should be confronted with balanced lists: (a) if lists are not sufficiently parallel, the variance of within-subjects effect sizes due to the experimental manipulation will be inflated, reducing statistical power---if there is no proper statistical control of the stimulus materials, which is often the case [@judd2012]; (b) highly unbalanced sets may introduce secondary effects beyond the influence of the intrinsic stimulus features. For example, highly biased lists may induce a strategic behavior shift between experimental blocks, possibly due to suspicion, boredom or frustration. These effects invalidate any observed differences between conditions and importantly, this bias cannot be accounted for via statistical techniques.

Traditionally, assigning stimuli to experimental conditions was done manually [e.g., @schaper2019metamory] or randomly [e.g., @kroneisen2018], both of which methods however do not ensure that the stimulus sets will be similar [@papenberg2020; @cutler1981; @vancasteren2007; @armstrong2012]. Generally, automated approaches are preferred that quantify similarity on the basis of the stimulus features and use an algorithm to assign stimuli in such a way that this similarity is optimized [@brusco2019; @papenberg2020]. One approach that has recently gained traction

**Using anticlustering to equate stimulus sets**

**BUT DISPERSION IS ALSO IMPORTANT!**

For example, in tasks of recognition memory, participants commonly have to decide if a target stimuli (e.g., a word) has been presented in an earlier phase of the experiment. In this case, words that are orthographically similar to a target word should serve as distractors.

The BILS algorithm has been available in the anticlust package since 2021, but we assume that many researchers do not yet know about its potential, also given that the introduction by @brusco2019 was rather technical in nature. For now, users seem to rely on the unicriterion methods focusing on between-set similarity only. Therefore, we provide readers with an application-focused introduction and extend BILS with parameters that may be useful when designing experiments.

# Discussion

\newpage

# References

\begingroup
<div id="refs" custom-style="Bibliography"></div>
\endgroup